Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalogue Data

### 2012-11-22 doc/paper-lcr.tex
more draft editing

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight.
Such massive structures contribute an ``external convergence'' $\kappax$ that
in time delay gravitational lenses can dominate the uncertainty in the inferred distances.  We
characterise this uncertainty by marginalising over a probability density
function for $\kappax$: here, we investigate the use of a simple halo model to
estimate $\pr(\kappax|\data)$ given noisy estimates of the photometric redshift
and stellar mass of galaxies observed in optical imaging surveys of the field
in question. We use mock catalogues from the Millennium Simulation, and compare
our reconstructed $\pr(\kappax|\data)$ to ray-traced $\kappax$ truth values: 
taking into account realistic uncertainties on redshift and stellar masses, we
typically find that the reconstruction yields a factor of 2 improvement in precision of
over the global $\pr(\kappax)$. For any given line of sight the
convergence estimate can be made before the investment of expensive follow-up
observations. We find that the lowest $\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$ after applying our reconstruction. We find that selecting systems by the precision with which
$\kappax$ can be estimated leads to a sample of unbiased
time delay distance estimates with 1\% uncertainties due to external convergence effects.
Photometric data alone is sufficient to pre-select the best lenses and can be done
before investment in lightcurve monitoring. Preferentially selecting lines of
sight with high external shear (as might happen in an assembled sample of
quadruply-imaged time delay lenses) produces time-delay distance estimates with
typical uncertainties of 1.5\% and biases of 0.3\%. We suggest areas for the improvement of this general
analysis framework that should allow yet more accurate cosmological inferences
to be made.

@@

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012b}.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight the
Green et al. results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense lines.

@@

This paper is organized as follows. We review the relevant gravitational
lensing theory in~\Sref{sec:theory}, and the Millennium Simulation
ray-traced convergences and mock catalogues in \Sref{sec:MS}, before
introducing our simple reconstruction  model in \Sref{sec:model}. We
then test this model in two phases: first, in \Sref{sec:knownMh+z}, with
known redshift and halo mass for every galaxy in a lens field, in order
to quantify the irreducible uncertainty due to unseen mass, and second,
in \Sref{sec:obsMstar+z}, with realistic observational uncertainties on
the observed galaxies' stellar masses and redshifts. In
\Sref{sec:biases} we investigate the potential systematic error induced
by selecting a subset of sightlines. We discuss our results in
\Sref{sec:discuss} before concluding in \Sref{sec:conclude}.

@@

Mass distributions that are
physically associated with the lens galaxy will affect the stellar dynamics of
the lens galaxy, so that dynamical observations can break the internal
mass-sheet degeneracy by providing an additional estimate of the lens galaxy's
mass \citep[e.g.,][]{SuyuEtal2010}. In this work our aim is to quantify the
lensing effect of massive objects, regardless of their position along the line
of sight.

@@

To define a test line of sight, we draw a random sky position from
within the convergence maps, bilinearly interpolate between the pixels
of the map, and store the result as the ``true'' convergence for that
line of sight, $\kappax^{\rm true}$.

@@

Each halo in our catalogue then contributes to
the line of sight convergence by
\be
\label{eq:kappai}
\kappa_i =\Sigma_{i}/\Sigma_{\mathrm{cr}}(z_i,\zs),
\ee
where the critical surface density was defined in \Eref{eq:sigcrit}.
Following the first order approximation of \citet{HilbertEtal2009}
outlined in \Sref{sec:MS} above, we compute 
the total convergence from all the halos along the line
of sight using
\be 
\label{eq:kappasummu}
\kappah = \sum_{i} \kappa_i.
\ee

To estimate a halo's contribution to $\kappah$ we must first know the
halo's mass. For the sightlines of the \MS, halo masses are
known, but for  real lines of sight the halo masses have to be \infered
from whatever data is available. 
In this work we investigate the use of the empirical stellar mass --
halo mass relation for this purpose. The primary advantage of this
approach is that it is a simple, ``one size fits all'' relation that can
be applied regardless of the stellar mass observed. We discuss
additional observables that could be used to improve the precision of
the halo mass inference in \Sref{sec:discuss} below.

@@

This PDF contains a hidden assumption of an uninformative prior PDF for
$\kappah$ -- in the case of infinitely poorly measured stellar masses and
redshifts, $\pr(\kappah|\data)$ will have very long tails corresponding to
very overdense sightlines that we do not believe exist. The resolution of this
is to divide out the effective prior PDF that was applied during the halo mass
estimation process (which is broad, and hence close to uniform), and apply an
additional prior on $\kappah$, given by the global underlying $\kappah$
distribution from the simulation. In the limiting  case of very poor
photometric data, $\pr(\kappah|\data)$ then defaults to this distribution.

@@

We  take a large number of
calibration sightlines from the \MS catalogues and generate a mock
stellar mass catalogue for each one. We then reconstruct each lightcone in the
same way as we would an observed field,
estimate $\pr(\kappah|\data)$, and extract its median. Accumulating
these samples and their corresponding true values of $\kappax$, we can form the 
conditional distribution $\pr(\kappax|\kappah^{\rm med})$. 
Then, when inferring $\kappax$ from an observed photometric catalogue, we 
estimate $\pr(\kappah|\data)$ 
using the halo model, compute the median value 
$\kappa_{\rm h,obs}^{\rm med}$, and then use an alternative
version of 
\Eref{eq:kappaconv} to infer $\kappax$: 
\bea
\Pr(\kappax|\kappah^{\rm med},\mathcal{D}) &=& \int \dee\kappah^{\rm med} 
   \Pr(\kappax|\kappah^{\rm med},\mathcal{D}) \Pr(\kappah^{\rm med}|\mathcal{D}) \notag \\
\label{eq:calkappaconv}   
\eea
Because we only have a finite number of calibration lines of sight, we use all calibration
lines of sight with $kappah_{\rm calibration}^{\rm med} = kappah^{\rm med}\pm0.003$ to form
$\Pr(\kappax|\kappah^{\rm med},\mathcal{D})$, with each calibration line weighted
equally.

The procedure of reducing the full PDF for $\kappah$ to its median increases the 
importance of the simulation choice: the simulation needs to be a similar as possible
to the real universe or there is a possibility of systematic errors. The reader
should keep in mind that we have only performed this analysis using the \MS.

\section{Results: Perfect Halo Data}
\label{sec:knownMh+z}

@@

\begin{figure*}
\includegraphics[width=\textwidth]{figs/where_is_the_kappa.eps}
\caption[magcut]{Which objects dominate the external convergence for a line of
sight? From left to right, the three figures show the cumulative contribution
to the total  convergence from individual halos ($\kappah$) as a function of
1)  distance from the line of sight, 2) magnitude and 3) redshift.   In each
panel, the solid line is the median cumulative contribution over a large
sample of sightlines, while dashed and dot-dashed show the ranges enclosing 68
and 95 percent of the sightlines repectively.}
\label{fig:where}
\end{figure*}

@@

We now move on to consider the halo model reconstruction of line of sight mass
distributions given noisy astronomical observables. A typical imaging survey
can be expected to provide measurements of the positions and magnitudes of
galaxies in a field;  spectra for some of the objects may either come from a
synergistic survey, or from targeted follow-up. In this section we quantify
the errors induced by inferring the halo mass and redshift from these
observables.

As described in \Sref{sec:model:halos}, in this work we infer halo masses
given measurements of stellar mass. We will investigate two main sources of
error: inferring halo mass given observed stellar mass, and placing halos at
the wrong redshift due to photometric redshift error. Much work has already
focused on using photometric colours to infer stellar mass
\citep[\eg][]{AugerEtal2009} and redshifts \citep[\eg][]{BPZ}: we will
estimate the likely uncertainties on stellar mass and redshift based on this
work.

@@

While the \MS catalogues do already contain stellar masses for each galaxy, we
find that these are not well behaved at the low mass end due to unreasonably high dark matter stripping from satellite halos. We also wanted to be able to perform the functional test of
trying to recover the convergence having assumed the correct stellar
mass--halo mass relation. For these reasons, we assign a new stellar mass to
each halo in the \MS catalogues, according to the stellar mass--halo mass
relation of \citet{BehrooziEtal2010}. From this true  stellar mass we
simulate an observed stellar mass by drawing a sample from 
$\pr(\log{M^{*}_{\rm obs}}|\log{M^{*}_{\rm true}})$ which we take to be a
Gaussian of width $\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where
a spectroscopic redshift exists stellar masses can be estimated with typical
uncertainties of 0.15 dex \citep{AugerEtal2009}; however with photometric
redshifts stellar mass uncertainties are typically three times as large. We
use $\sigma_{M_*}=0.15$ dex for halos with a spectroscopic redshift and
$\sigma_{M_*}=0.45$ dex otherwise. For photometric redshift errors we draw a
redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take to be a Gaussian
of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where $z_{\rm spec}$
is the halo's true redshift in the \MS catalogue.

@@

Inferring the stellar mass of a galaxy from its magnitude and colours requires
an estimate of how far away the galaxy is; without a spectroscopic redshift
the \infered stellar mass is less precise. However, obtaining photometry has
much lower observational cost; over the next few years several large area
photometric surveys will reach 24th magnitude \citep{Euclid,LSST}, these will
provide sufficient data to reconstruct lines of sight {\it without} additional
observations. In principle the photometric redshift is correlated with the
\infered stellar mass, however we do not model this effect since the
convergence from the outskirts of an individual halo is only weakly \dependant
on redshift at fixed mass: the redshift error has only a small effect on the
inferred $\pr(\kappax|\data)$ when compared to the effect of uncertain stellar
masses.  With only photometric redshifts the uncertainty on
$\kappah$ is larger than in the spectroscopic case and this
\propogates into a broader $\pr(\kappax|\data)$, although the photometric
reconstruction still typically produces a 50\% improvement compared to the
precision of the global $\pr(\kappax)$ (\Fref{fig:widthsvsH}).

@@

\section{Systematic Errors due to Sample Selection}
\label{sec:bias}

@@

\subsubsection{Selecting random lines of sight}
\label{sec:bias:random}

@@

\subsubsection{Selecting only the lines of sight with tight $\pr(\kappax|\data)$}
\label{sec:bias:tightPDF}

@@

\subsubsection{Selecting only high shear lines of sight}
\label{sec:bias:tightPDF}

@@

By selecting only lines of sight with an external shear of $\gammax>0.03$ we
can test the impact of shear selection. We find that in this sample, $\kappax$
is systematically underestimated at the $\sim$0.003 level (corresponding to a
0.3\% systematic error in distance; in the absence of other sources of uncertainty, this
systematic error would be significant in a sample of 40 lenses). 
This effect is mitigated if only
the lines of sight with both $\gammax>0.03$ and well constrained $\pr(\kappax|\data)$ are
used, but biases at the $\sim$0.001 level remain. Since our model does not
include shear constraints it is not surprising that a selection function based on shear
can induce a bias. A more sophisticated model that includes the shear
recovered from the lens modelling may be less susceptible to this bias. The
toy selection model used here ($\gammax>0.03$) is likely to be a much
stronger selection than reality, but nevertheless it is worth noting that the
reconstruction \proceedure will be biased if lens selection functions are
extreme and unaccounted for.

@@

We have shown that reconstructing the matter due to halos along any line of
sight can improve the constraints on the external convergence along that line
of sight, with, depending on sample selection,  the width of
$\pr(\kappax|\data)$ being up to a factor of two smaller than the global
$\pr(\kappax)$, and  we have quantified the residual bias in the $\kappax$
estimates under our calibration assumptions. We now revisit these assumptions.

The total convergence along a line of sight is strongly correlated with the
reconstructed $\kappah$. However, since our model ignores voids and assumes
all halos follow a spherical truncated-NFW profile our halo model does not
include all of the relevant physics; hence, the width of our resulting
$\pr(\kappax|\data)$ is still typically $\sim$0.01 for any given lightcone,
even with a perfect knowledge of every halo's virial mass and redshift. To
make further progress a more advanced treatment of both voids and halos will
be necessary. \citet{KarpenkaEtal2012} find that their supernova data prefer,
in the context of their simple halo model, a truncated isothermal profile for
the galaxy mass distributions. Weak galaxy-galaxy and cluster-galaxy  lensing
studies \citep[\eg][]{GavazziEtal2006,JohnstonEtal2007} also suggest an
approximately isothermal profile for the mass-shear correlation function;
however, at large radii these profiles are dominated by the effects of
neighbouring halos, which should be taken into account by our model. Further
iterations on the analysis presented here should include the stellar mass
distributions that cause the isothermal density profiles in the centres of
galaxies, although we do not expect this to have a big impact on the model.
Likewise,  halo ellipticity may have some small effect on the predicted
convergence.

@@

Interestingly, we have found that it is the most empty lines of sight that 
can be reconstructed with the highest precision. $\kappax$ estimates for empty
lines of sight receive few contributions from halos, and a large contribution
from voids; since they have the tightest PDFs after the reconstruction it
seems that our model's biggest uncertainties are driven by naively
reconstructing halos rather than neglecting voids. With a photometric
reconstruction of the field there is a significant broadening, but the
resulting precision is still typically 50\%  higher than using the global
$\kappax$ distribution.

Since the uncertainty on $\pr(\kappax|\data)$ is $\sim$0.01 even
with perfect knowledge of halo mass and redshift, it seems that modelling uncertainties
from voids, deviations from sphericity and dark matter clumping within the
main halo dominate the total $\kappax$ uncertainty even with very
uncertain stellar masses. Inferring halo ellipticity and dark matter
clumping will likely always remain a difficulty for line of sight
reconstruction; but as \Fref{fig:magcut} shows, observing deeper
than 24th magnitude does not significantly help the reconstruction.

For a small fraction of lines of sight, $\pr(\kappax|\data)$ remains very
broad even after applying our reconstruction; these lines of sight are
typically the most over-dense lines of sight in the universe. For
time-delay cosmography the most observationally expensive tasks are the
light-curve monitoring and high resolution follow-up imaging and spectroscopy,
while making photometric observations of a
4$\times$4 arcminute patch of sky survey down to 24th magnitude is
relatively cheap. We have shown that with a single epoch observation of
the region around a strong lens it is possible to infer $\pr(\kappax|\data)$.
If a line of sight has a broad $\pr(\kappax|\data)$ it can be rejected {\it
before} the investment of follow-up. We find that
selecting only the lines of sight with the most tightly constrained
$\pr(\kappax)$ does not induce a bias in $\kappax$, and hence distance.
For the most over-dense lines of sight,
the precision of $\pr(\kappax|\data)$ is improved noticeably by the addition
of spectroscopic information; currently insufficient time-delay systems are
know to choose a subset based on their reconstructed $\pr(\kappax|\data)$.
Spectroscopic coverage will definitely be useful for extracting cosmology from
over-dense lines of sight.

The method we have outlined can also be used to estimate the external shear
along a line of sight. Shear is an observable that can be extracted from
strong lens modelling, however there is a degeneracy between internal and
external shear. When the Einstein Ring data are very good it is possible to
disentangle external and internal shear \citep[\eg][]{xxx}, but there are
still significant uncertainties: \citet{WongEtal2011} attempted to match the
shear from strong lens models with a reconstruction of the local lens group
environment, but found a tension between the strong lens model and the
reconstruction of the environment. Given the \citet{WongEtal2011} results
it is unclear whether the external shear from lens models can be
reconciled with a line of sight reconstruction. Alternatively, it may be possible 
to infer external shear using weak lensing information from near the line of sight. 
If the true external shear can be
measured, it provides an additional constraint on which of the \MS lines of
sight are similar to the reconstructed line of sight. \citet{SuyuEtal2012}
found that for RXJ1131 combining shear constraints with galaxy number count
over-density, gave a significantly different $\pr(\kappax|\gamma,N_{45})$
compared to the PDF from number count over-density alone,
$\pr(\kappax|N_{45})$.

>>>>

High precision cosmological distance measurements towards individual objects
such as time delay gravitational lenses or type Ia supernovae are affected by
weak lensing perturbations by galaxies and groups along the line of sight.
Such massive structures contribute an ``external convergence'' $\kappax$ that
in time delay gravitational lenses can dominate the uncertainty in the inferred distances.  We
characterise this uncertainty by marginalising over a probability density
function for $\kappax$: here, we investigate the use of a simple halo model to
estimate $\pr(\kappax)$ given noisy estimates of the photometric redshift
and stellar mass of galaxies observed in optical imaging surveys of the field
in question. We use mock catalogues from the Millennium Simulation, and compare
our reconstructed $\pr(\kappax|\data)$ to ray-traced $\kappax$ truth values: 
taking into account realistic uncertainties on redshift and stellar masses, we
typically find that the reconstruction yields a factor of 2 improvement in precision of
over the global $\pr(\kappax)$. For any given line of sight the
convergence estimate can be made before the investment of expensive follow-up
observations. We find that the lowest $\kappax$ lines of sight have the best constrained 
$\pr(\kappax)$ after applying our reconstruction. We find that selecting systems by the precision with which
$\kappax$ can be estimated leads to a sample of unbiased
time delay distance estimates with 1\% uncertainties due to external convergence effects.
Photometric data alone is sufficient to pre-select the best lenses and can be done
before investment in lightcurve monitoring. Preferentially selecting lines of
sight with high external shear (as might happen in an assembled sample of
quadruply-imaged time delay lenses) produces time-delay distance estimates with
typical uncertainties of 1.8\% and biases of 0.8\%. We suggest areas for the improvement of this general
analysis framework that should allow yet more accurate cosmological inferences
to be made.

@@

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuetAl2012b}.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight the
Green et al. results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense lines.

@@

This paper is organized as follows. We review the relevant gravitational
lensing theory in~\Sref{sec:theory}, and the Millennium Simulation
ray-traced convergences and mock catalogues in \Sref{sec:MS}, before
introducing our simple reconstruction  model in \Sref{sec:model}. We
then test this model in two phases: first, in \Sref{sec:knownMh+z}, with
known redshift and halo mass for every galaxy in a lens field, in order
to quantify the irreducible uncertainty due to unseen mass, and second,
in \Sref{sec:obsMstar+z}, with realistic observational uncertainties on
the observed galaxies' stellar masses and redshifts. In
\Sref{sec:biases} we investigate the potential systematic error induced
by selecting a subset of lines of sight. We discuss our results in
\Sref{sec:discuss} before concluding in \Sref{sec:conclude}.

@@

Mass distributions that are
physically associated with the lens galaxy will affect the stellar dynamics of
the lens galaxy, so that dynamical observations can break the internal
mass-sheet degeneracy by providing an additional estimate of the lens galaxy's
mass \citep[e.g.,][]{Koopmans2004;SuyuEtal2010}. In this work our aim is to quantify the
lensing effect of massive objects, regardless of their position along the line
of sight.

@@

To define a test line of sight, we draw a random sky position from
within the convergence maps, bi-linearly interpolate between the pixels
of the map, and store the result as the ``true'' convergence for that
line of sight, $\kappax^{\rm true}$.

@@

Each halo in our catalogue then contributes to
the line of sight convergence by
\be
\label{eq:kappai}
\kappa_i =\Sigma_{i}/\Sigma_{\mathrm{cr}}(z_i,\zs),
\ee
where $\Sigma_{\mathrm{cr}}(z_i,\zs)$ is the critical surface density \comments{was defined in \Eref{eq:sigcrit}}.
Following the first order approximation of \citet{HilbertEtal2009}
outlined in \Sref{sec:MS} above, we compute 
the total convergence from all the halos along the line
of sight using
\be 
\label{eq:kappasummu}
\kappah = \sum_{i} \kappa_i.
\ee

To estimate a halo's contribution to $\kappah$ we must first know the
halo's mass. For the lines of sight of the \MS, halo masses are
known, but for  real lines of sight the halo masses have to be \infered
from whatever data is available. 
In this work we investigate the use of the empirical stellar mass --
halo mass relation for this purpose. The primary advantage of this
approach is that it is a simple, ``one size fits all'' relation that can
be applied regardless of the stellar mass observed. We discuss
additional observables that could be used to improve the precision of
the halo mass inference in \Sref{sec:discuss} below.

@@

This PDF contains a hidden assumption of an uninformative prior PDF for
$\kappah$ -- in the case of infinitely poorly measured stellar masses and
redshifts, $\pr(\kappah|\data)$ will have very long tails corresponding to
very over-dense lines of sight that we do not believe exist. The resolution of this
is to divide out the effective prior PDF that was applied during the halo mass
estimation process (which is broad, and hence close to uniform), and apply an
additional prior on $\kappah$, given by the global underlying $\kappah$
distribution from the simulation. In the limiting  case of very poor
photometric data, $\pr(\kappah|\data)$ then defaults to this distribution.

@@

We  take a large number of
calibration lines of sight from the \MS catalogues and generate a mock
stellar mass catalogue for each one. We then reconstruct each lightcone in the
same way as we would an observed field,
estimate $\pr(\kappah|\data)$, and extract its median. Accumulating
these samples and their corresponding true values of $\kappax$, we can form the 
conditional distribution $\pr(\kappax|\kappah^{\rm med})$. 
Then, when inferring $\kappax$ from an observed photometric catalogue, we 
estimate $\pr(\kappah|\data)$ 
using the halo model, compute the median value 
$\kappa_{\rm h,obs}^{\rm med}$, and then use an alternative
version of 
\Eref{eq:kappaconv} to infer $\kappax$: 
\bea
\Pr(\kappax|\kappah^{\rm med},\mathcal{D}) &=& \int \dee\kappah^{\rm med} 
   \Pr(\kappax|\kappah^{\rm med},\mathcal{D}) \Pr(\kappah^{\rm med}|\mathcal{D}) \notag \\
\label{eq:calkappaconv}   
\eea
Because we only have a finite number of calibration lines of sight, we use all calibration
lines of sight with $kappah_{\rm calibration}^{\rm med} = kappah^{\rm med}\pm0.003$ to form
$\Pr(\kappax|\kappah^{\rm med},\mathcal{D})$, with each calibration line weighted
equally.

The procedure of reducing the full PDF for $\kappah$ to its median increases the 
importance of the simulation choice: the simulation needs to be a similar as possible
to the real universe or there is a possibility of systematic errors. 
\section{Results: Perfect Halo Data}
\label{sec:knownMh+z}

@@

\begin{figure*}
\includegraphics[width=\textwidth]{figs/where_is_the_kappa.eps}
\caption[magcut]{Which objects dominate the external convergence for a line of
sight? From left to right, the three figures show the cumulative contribution
to the total  convergence from individual halos ($\kappah$) as a function of
1)  distance from the line of sight, 2) magnitude and 3) redshift.   In each
panel, the solid line is the median cumulative contribution over a large
sample of lines of sight, while dashed and dot-dashed show the ranges enclosing 68
and 95 percent of the lines of sight respectively.}
\label{fig:where}
\end{figure*}

@@

We now move on to consider the halo model reconstruction of line of sight mass
distributions given noisy astronomical observables. A typical imaging survey
can be expected to provide measurements of the positions and magnitudes of
galaxies in a field;  spectra for some of the objects may either come from a
synergistic survey, or from targeted follow-up. In this section we quantify
the uncertainties induced by inferring the halo mass and redshift from these
observables.

As described in \Sref{sec:model:halos}, in this work we infer halo masses
given measurements of stellar mass. We will investigate two main sources of
uncertainty: inferring halo mass given observed stellar mass, and placing halos at
the wrong redshift due to photometric redshift uncertainty. Much work has already
focused on using photometric colours to infer stellar mass
\citep[\eg][]{AugerEtal2009} and redshifts \citep[\eg][]{BPZ}: we will
estimate the likely uncertainties on stellar mass and redshift based on this
work.

@@

While the \MS catalogues do already contain stellar masses for each galaxy, we
find that these are not well behaved at the low mass end due to unreasonably high dark matter stripping from satellite halos. We also wanted to be able to perform the functional test of
trying to recover the convergence having assumed the correct stellar
mass--halo mass relation. For these reasons, we assign a new stellar mass to
each halo in the \MS catalogues, according to the stellar mass--halo mass
relation of \citet{BehrooziEtal2010}. From this true  stellar mass we
simulate an observed stellar mass by drawing a sample from 
$\pr(\log{M^{*}_{\rm obs}}|\log{M^{*}_{\rm true}})$ which we take to be a
Gaussian of width $\sigma_{M_*}$ centred on $\log(M*_{\mathrm {true}})$. Where
a spectroscopic redshift exists stellar masses can be estimated with typical
uncertainties of 0.15 dex \citep{AugerEtal2009}; however with photometric
redshifts stellar mass uncertainties are typically three times as large. We
use $\sigma_{M_*}=0.15$ dex for halos with a spectroscopic redshift and
$\sigma_{M_*}=0.45$ dex otherwise. For photometric redshift uncertainties we draw a
redshift from $\pr(z_{\rm true}|z_{\rm obs})$ which we take to be a Gaussian
of width $0.1(1+z_{\rm spec})$ centred on $z_{\rm spec}$, where $z_{\rm spec}$
is the halo's true redshift in the \MS catalogue.

@@

Inferring the stellar mass of a galaxy from its magnitude and colours requires
an estimate of how far away the galaxy is; without a spectroscopic redshift
the \infered stellar mass is less precise. However, obtaining photometry has
much lower observational cost; over the next few years several large area
photometric surveys will reach 24th magnitude \citep{Euclid,LSST}, these will
provide sufficient data to reconstruct lines of sight {\it without} additional
observations. In principle the photometric redshift is correlated with the
\infered stellar mass, however we do not model this effect since the
convergence from the outskirts of an individual halo is only weakly \dependant
on redshift at fixed mass: the redshift uncertainty has only a small effect on the
inferred $\pr(\kappax|\data)$ when compared to the effect of uncertain stellar
masses.  With only photometric redshifts the uncertainty on
$\kappah$ is larger than in the spectroscopic case and this
\propogates into a broader $\pr(\kappax|\data)$, although the photometric
reconstruction still typically produces a 50\% improvement compared to the
precision of the global $\pr(\kappax)$ (\Fref{fig:widthsvsH}).

@@

\section{Systematic Errors due to Sample Selection}
\label{sec:biases}

@@

\subsection{Selecting random lines of sight}
\label{sec:bias:random}

@@

\subsection{Selecting only the lines of sight with tight $\pr(\kappax|\data)$}
\label{sec:bias:tightPDF}

@@

\subsection{Selecting only high shear lines of sight}
\label{sec:bias:tightPDF}

@@

By selecting only lines of sight with an external shear of $\gammax>0.05$ we
can test the impact of shear selection. We find that in this sample, $\kappax$
is systematically underestimated at the $\sim$0.008 level (corresponding to a
0.8\% systematic error in distance); in the absence of other sources of uncertainty, this
systematic error would be significant in a sample of 5 lenses. Approximating strong-lens
modelling uncertainties as an unbiased gaussian with width 4\% on time-delay distance a 0.8\% bias would be
significant for $\sim$20 lenses. 
The systematic error is mitigated if only
the lines of sight with both $\gammax>0.05$ and well constrained $\pr(\kappax|\data)$ are
used, but biases at the $\sim$0.005 level remain. The
toy selection model used here ($\gammax>0.05$) is likely to be a much
stronger selection than reality, but nevertheless it is worth noting that the
reconstruction \proceedure will be biased if lens selection functions are
extreme and unaccounted for. Since our model does not
include shear constraints it is not surprising that a selection function based on shear
can induce a bias. A more sophisticated model that includes the shear
recovered from the lens modelling may be less susceptible to this bias.

The halo model can also be used to estimate the external shear
along a line of sight. Shear is an observable that can be extracted from
strong lens modelling, however there is a degeneracy between internal and
external shear. When the Einstein Ring data are very good it is possible to
disentangle external and internal shear \citep[\eg][]{SuyuEtal2010}, but there are
still significant uncertainties: \citet{WongEtal2011} attempted to match the
shear from strong lens models with a reconstruction of the local lens group
environment, but found a tension between the strong lens model and the
reconstruction of the environment. Given the \citet{WongEtal2011} results
it is unclear whether the external shear from lens models can be
reconciled with a line of sight reconstruction. Alternatively, it may be possible 
to infer external shear using weak lensing information from near the line of sight. 
If the true external shear can be
measured, it provides an additional constraint on which of the \MS lines of
sight are similar to the reconstructed line of sight. \citet{SuyuEtal2012}
found that for RXJ1131 combining shear constraints with galaxy number count
over-density, gave a significantly different $\pr(\kappax|\gamma,N_{45})$
compared to the PDF from number count over-density alone,
$\pr(\kappax|N_{45})$.

@@

\comments{
We have shown that reconstructing the matter due to halos along any line of
sight can improve the constraints on the external convergence along that line
of sight, with, depending on sample selection,  the width of
$\pr(\kappax|\data)$ being up to a factor of two smaller than the global
$\pr(\kappax)$, and  we have quantified the residual bias in the $\kappax$
estimates under our calibration assumptions. We now revisit these assumptions.
}

The total convergence along a line of sight is strongly correlated with the
reconstructed $\kappah$. However, since our model ignores voids and assumes
all halos follow a spherical truncated-NFW profile our halo model does not
include all of the relevant physics; hence, the width of our resulting
$\pr(\kappax|\data)$ is still typically $\sim$0.01 for any given lightcone,
even with a perfect knowledge of every halo's virial mass and redshift. To
make further progress a more advanced treatment of both voids and halos will
be necessary. \citet{KarpenkaEtal2012} find that their supernova data prefer,
in the context of their simple halo model, a truncated isothermal profile for
the galaxy mass distributions. Galaxy-galaxy and cluster-galaxy  lensing
studies \citep[\eg][]{GavazziEtal2007,JohnstonEtal2007} also suggest an
approximately isothermal profile for the mass-shear correlation function;
however, at large radii these profiles are dominated by the effects of
neighbouring halos, which should be taken into account by our model. Further
iterations on the analysis presented here should include the stellar mass
distributions that cause the isothermal density profiles in the centres of
galaxies, although we do not expect this to have a big impact on the model.
Likewise,  halo ellipticity may have some small effect on the predicted
convergence. It is possible that baryonic physics may alter   
the radial profile of dark matter halos, even on large scales. \citet{Semboloni+2012}
found that baryonic physics had a 30-40\% impact on three-point shear statistics, it is
unclear how baryons affect one-point convergence statistics compared to the pure dark matter
simulations used in this work.

@@

\comments{Interestingly, we have found that it is the most empty lines of sight that 
can be reconstructed with the highest precision. $\kappax$ estimates for empty
lines of sight receive few contributions from halos, and a large contribution
from voids; since they have the tightest PDFs after the reconstruction it
seems that our model's biggest uncertainties are driven by naively
reconstructing halos rather than neglecting voids. With a photometric
reconstruction of the field there is a significant broadening, but the
resulting precision is still typically 50\%  higher than using the global
$\kappax$ distribution.

Since the uncertainty on $\pr(\kappax|\data)$ is $\sim$0.01 even
with perfect knowledge of halo mass and redshift, it seems that modelling uncertainties
from voids, deviations from sphericity and dark matter clumping within the
main halo dominate the total $\kappax$ uncertainty even with very
uncertain stellar masses. Inferring halo ellipticity and dark matter
clumping will likely always remain a difficulty for line of sight
reconstruction; but as \Fref{fig:magcut} shows, observing deeper
than 24th magnitude does not significantly help the reconstruction.
}

\comments{For a small fraction of lines of sight, $\pr(\kappax|\data)$ remains very
broad even after applying our reconstruction; these lines of sight are
typically the most over-dense lines of sight in the universe. For
time-delay cosmography the most observationally expensive tasks are the
light-curve monitoring and high resolution follow-up imaging and spectroscopy,
while making photometric observations of a
4$\times$4 arcminute patch of sky survey down to 24th magnitude is
relatively cheap. We have shown that with a single epoch observation of
the region around a strong lens it is possible to infer $\pr(\kappax|\data)$.
If a line of sight has a broad $\pr(\kappax|\data)$ it can be rejected {\it
before} the investment of follow-up. We find that
selecting only the lines of sight with the most tightly constrained
$\pr(\kappax)$ does not induce a bias in $\kappax$, and hence distance.
For the most over-dense lines of sight,
the precision of $\pr(\kappax|\data)$ is improved noticeably by the addition
of spectroscopic information; currently insufficient time-delay systems are
know to choose a subset based on their reconstructed $\pr(\kappax|\data)$.
Spectroscopic coverage will definitely be useful for extracting cosmology from
over-dense lines of sight.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\comments{%
\begin{figure}
\includegraphics[width=\columnwidth]{figs/biasplots3.eps}
\caption{Convergence estimation accuracy using the full  $\pr(\kappax|\kappah)$
distribution, rather than its median.  As in \Fref{fig:biasplots}, solid,
dashed and dotted lines show combinations of 20, 5 and 2 lines of sight
respectively, while green lines show the inferences 
from photometry alone, red lines from photometry and targeted spectroscopy 
(assuming a spectroscopic redshift for all
halos with $i<23$ and all halos within 1 arcminute and $i<24$), and blue lines
from the unrealistic assumption of a spectroscopic redshift for all galaxies
in the field. }
\label{fig:no-median-tricks}
\end{figure}
}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The reduction of $\pr(\kappah|\data)$ to just its median value, before using
it to look up the appropriate $\pr(\kappax)$ from the simulated sightline
ensemble, is reminiscent of the procedure followed by \citet{SuyuEtal2010},
and Greene et al (in prep.) The reconstruction method can be seen as the
limiting case of the weightings explored by Greene et al: rather than using
weighted number counts in an aperture, we treat each galaxy individually, and
predict  $\kappa_{\rm h,obs}^{\rm med}$ for use in their place.

\comments{ One
significant disadvantage of these types of method is that they impose a
bottleneck in the flow of information. Once you have compressed to one number,
it is difficult to incorporate more information later; you have to recompute
all the calibrations. Using the full PDF for $\kappah$ is the more correct
thing to do: at present, however, it suffers from significant bias in the
photometric data case. \Fref{fig:no-median-tricks} shows how $\kappax$ is
overestimated, as a result of the high scatter in halo mass for objects with 
high stellar mass. As has already been noted, 
future work should focus on increasing the precision of the halo
mass estimates. 
}
